\begin{}
    \section*{\centering REFERENCES}
    [1] Holmstrom Mark, Dylan Liu and Christopher Vo, "Machine learning applied to weather forecasting", Meteorol. Appl, pp. 1-5, 2016.


\noindent[2] Sunita B. Aher and L. M. R. J. Lobo, "Comparative study of classification algorithms", International Journal of Information Technology, vol. 5.2, pp. 239-243, 2012.

\noindent[3] Sharma, S.; Sharma, S. Activation functions in neural networks. Towards Data Sci.no. 6,pp.310–316,2017


\noindent[4] Nalluri Sravani, Ramasubbareddy Somula and G Kannayaram, "Weather Prediction Using Clustering Strategies in Machine Learning", Journal of Computational and Theoretical Nanoscience, vol. 16, no. 5, pp. 1977-1981, 2019.

\noindent[5] S. D Drobot, M. Chapman, P. A. Pisano and B. B. McKeever, "Using vehicles as mobile weather platforms" in Data and Mobility: Transforming Information into Intelligent Traffic and Transportation Services, Springer, pp. 203-214, 2010.

\noindent [6]S. B. Lajevardi and B. Minaei-Bidgoli, "Combination of Time Series Decision Tree and Clustering : A Case Study in Aerology Event Prediction", International Conference on Computer and Electrical Engineering, pp. 111-115, 2008.

\noindent[7] D. B. Percival and A.T. Walden, "Methods for time series analysis" in , Cambridge, MA:Cambridge University Press, 2000.

\noindent[8] S. Shaminder, B. Pankaj and G. Jasmeen, "Time Series based Temperature Prediction using Back Propagation with Genetic Algorithm Technique", International Journal of Computer Science, vol. 8, pp. 28-32, 2011.

\noindent[9]   Imran Maqsood, Ajith Muhammad Riaz Khan, Abraham
An ensemble of neural network for weather forecasting, volume 13, pp 112 - 122,2004.

\noindent[10] Amanpreet Kaur, Sharma
Artificial neural networks in forecasting maximum and minimum relative humidity
International Journal of Computer Science and Network Security, volume 11, issue 5, p. 197 - 199,2011.

\noindent[11] G Vamsi Krishna
An integrated approach for weather forecasting based on data mining and forecasting analysis
International Journal of Computer Applications, volume 120, issue 11, p. 975 - 8887, 2015.

\noindent[12] Neeraj Kumar, Govind Kumar Jha
A time series ANN approach for weather forecasting
International Journal of Control Theory and Computer Modeling (IJCTCM), volume 3, 2013.

\noindent[13] N. Hasan, M. T. Uddin and N. K. Chowdhury, "Automated weather event analysis with machine learning", Proc. IEEE 2016 International Conference on Innovations in Science Engineering and Technology (ICISET), pp. 1-5, 2016.

\noindent[14] A. G. Salman, B. Kanigoro and Y. Heryadi, "Weather forecasting using deep learning techniques", Proc. IEEE 2015 International Conference on Advanced Computer Science and Information Systems (ICACSIS), pp. 281-285, 2015.

\noindent[15]L. L. Lai, H. Braun, Q. P. Zhang, Q. Wu, Y. N. Ma, W. C. Sun, et al., "Intelligent weather forecast", Proc. IEEE 2004 International Conference on Machine Learning and Cybernetics, pp. 4216-4221, 2004.

\noindent [16]Bashar Muneer Yahya and Dursun Zafer Seker, "Designing Weather Forecasting Model Using Computational Intelligence Tools", Applied Artificial Intelligence, vol. 33:2, pp. 137-151, 2019.

\noindent[17]Vladimir M. Krasnopolsky and Michael S. FoxRabinovitz, "Complex hybrid models combining deterministic and machine learning components for numerical climate modeling and weather prediction", Neural Networks19, vol. 2, pp. 122-134, 2006.

\noindent[18] S. D Drobot, M. Chapman, P. A. Pisano and B. B. McKeever, "Using vehicles as mobile weather platforms" in Data and Mobility: Transforming Information into Intelligent Traffic and Transportation Services, Springer, pp. 203-214, 2010.

\noindent [19] Predicting Wind Speed, Energies, 14, 20, (M. Wiston and M. KM, "Weather Forecasting: From the Early Weather Wizards to Modern-day Weather Predictions", Journal of Climatology & Weather Forecasting, vol. 6, pp. 1-9, 2018.

\noindent[20]M. Biswas, T. Dhoom and S. Barna, "Weather Forecast Prediction: An Integrated Approach for Analyzing and Measuring Weather", Data International Journal of Computer Applications, vol. 182, no. 4, pp. 20-24, 2018.
\end{}



