% Acknowledgements
\addcontentsline{toc}{chapter}{ACKNOWLEDGEMENT}
\section*{\centering{\normalsize ACKNOWLEDGEMENT}}
\noindent We would like to express our deep sense of gratitude to our supervisor and guide Dr. Atul Kumar Dwivedi, Department of Communication Engineering, BIET, Jhansi, for his constant guidance and motivation during the project work from last one year. We really appreciate and value their  guidance and constant motivation from the beginning of the project. Without his guidance and persistent help this project would not have been possible. 
We are also indebted to Professor Deepak Nagaria, Head Department of Electronics and Communications Engineering for his encouraging co-operation and continued interest in shaping this project.
We express our sincere gratitude to Prof. P. M. Pandey, Director, BIET, Jhansi, who has been instrumental in providing an ideal environment for wholesome individual development and appreciating talents in both academic and extracurricular activities.
We extend our sincere thanks to all faculties of the Department of Electronics and communication engineering, BIET, Jhansi, for their focused guidance and encouragement.
We would like to thank our parents and friends and all those people who have supported me directly or indirectly to complete this project work. 

\par	
	\vspace{0.008cm}
	\noindent
		\begin{tabular}{lcr}
		\\
		\bfseries  Ayush Pandey & \hspace{1.5cm} \textbf{Mushab Rehman} & \bfseries \hspace{2.5cm}  Vipin Kumar\\
			Roll no. 1904331017 & \hspace{1.5cm}Roll no. 1904331035  &
   Roll no.1904331065
		\end{tabular}